---
title: "Bennett Creek hydrology updates"
author: "Megan Sears"
date: "`r Sys.Date()`"
output: rmdformats::downcute
editor_options:
  chunk_output_type: console
---

<style type="text/css">

body, td {
   font-size: 25px;
}
code.r{
  font-size: 25px;
}
pre {
  font-size: 25px
}
</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      fig.width = 12,
                      fig.height = 6)

library(tidyverse)
library(lubridate)
library(raster)
library(rgdal)
library(sf)
library(sp)
library(mapview)
library(terra)
library(tmap)
library(readr)
library(gridExtra)
library(ggplot2); theme_set(theme_bw(base_size = 16))
library(here)
library(irr)
library(plotly)
library(kableExtra)
library(rmdformats)
library(shinyjs)

```

# Sites and sensors

```{r spatial}

tmap_mode('view')

#bennett polygons
ben <- vect('GIS/mulch_study_watersheds.shp') %>%
  terra::project(., 'EPSG:26913')

ben <- as(ben, "Spatial")

sensors <- read_csv('GIS/benn_sensors.csv') %>%
  st_as_sf(coords = c('x', 'y'), crs = 4326)

tm_shape(ben) +
  tm_borders(col = 'black') +
  tm_text('ID') +
  tm_shape(sensors) +
  tm_symbols(col = 'Type')


# cpf polygon - only  bring in once
cpf <- vect('GIS/cameron_boundary.shp') %>%
   terra::project(., 'EPSG:26913')

r <- rast('E:/MRMS_2022/RadarOnly_QPE_01H_00.00_20220716-000000.grib2') %>%
  terra::project(., 'EPSG:26913')

#crop raster to cpf
crop <- terra::crop(r, ext(cpf))

names(crop) <- 'p_mmhr'

```

# MRMS overview

* "MRMS is a system with fully-automated algorithms that quickly and intelligently integrate data streams from multiple radars, surface and upper air observations, lightning detection systems, satellite observations, and forecast models. Numerous two-dimensional multiple-sensor products offer assistance for hail, wind, tornado, quantitative precipitation estimations, convection, icing, and turbulence diagnosis." Information found here: https://www.nssl.noaa.gov/projects/mrms/
* In this area, it only includes the radar data (not integrated with forecast models, etc.)
* 1-km spatial resolution
* Hourly rainfall data

**MRMS rainfall data for 2022-07-15 at 17:00**

```{r spatial mrms example}

tm_shape(crop) +
  tm_raster(palette = 'Blues') +
  tm_shape(ben) +
  tm_borders(col = 'black') +
  tm_text('ID') +
  tm_shape(sensors) +
  tm_symbols(col = 'Type', size = 0.01)

```


```{r load data}

#bring in tb and sum as hourly
tip <- read_csv('bennett_rain_TB.csv') %>% 
  mutate(hour = hour(datetime),
         datetime = ymd_hm(datetime),
         date = as.Date(datetime)) %>%
  group_by(site, date, hour) %>%
  summarize(hr_p_mm = sum(precip_mm)) %>%
  mutate(datetime = ymd_h(paste(date, hour))) %>%
  filter(datetime > '2021-07-19 23:00') %>%
  ungroup() %>%
  dplyr::select(-c(date, hour)) %>%
  rename(p_mmhr = hr_p_mm) %>%
  mutate(source = 'tb')

#bring in mrms hourly  
mrms21 <- read_csv('mrms_21.csv') %>%
  dplyr::select(-c(...1, ID)) %>%
  rename(site = name) %>%
  mutate(site = tolower(site)) %>%
  mutate(datetime  = datetime - (2 * 60 * 60)) #should only be -1

mrms22 <- read_csv('mrms_22.csv') %>%
  rename(site = name) %>%
  mutate(site = tolower(site)) %>%
  mutate(datetime  = datetime - (2 * 60 * 60)) #should only be -1

mrms <- bind_rows(mrms21, mrms22) %>%
  mutate(source = 'mrms')

#bind together mrms and tb
all <- bind_rows(mrms, tip) %>%
  mutate(year = year(datetime)) %>%
  group_by(source, year, site) %>%
  mutate(cump = cumsum(p_mmhr)) %>%
  ungroup() %>%
  arrange(datetime)

```

# MRMS rainfall metrics

```{r mrms rainfall}

load('rain_metrics.RData')

site_id <- c('me', 'mm', 'mw', 'ue', 'um', 'uw')

get_radar_intens <- function(site_id){
  
mrms_prep <- mrms %>%
  filter(site == site_id,
         p_mmhr > 0) %>%
  arrange(datetime) %>%
  mutate(site == paste0(site_id))
  
mrms_intens <- get_setup(mrms_prep, mrms_prep$datetime) %>%
  group_by(event, site) %>%
  summarize(MI60 = max(p_mmhr),
            P = sum(p_mmhr),
            event = mean(event),
             duration = max(datetime) - min(datetime),
             endtime = max(datetime),
            starttime = min(datetime)) %>%
  mutate(duration_hr = as.numeric(duration/60/60))

return(mrms_intens)

}

mrms_mi60 <- site_id %>%
  map(~ get_radar_intens(site_id = .)) %>%
  bind_rows() %>%
  mutate(year = year(endtime)) %>%
  filter(P > 1)

mrms_mi60 %>%
ggplot(aes(site, MI60)) +
  geom_boxplot() +
  facet_wrap(~year) +
  ggtitle('MRMS MI60') +
  labs(y = 'MI60 (mm/hr)')

mrms_mi60 %>%
ggplot(aes(site, P)) +
  geom_boxplot() +
  facet_wrap(~year) +
  ggtitle('MRMS rainfall') +
  labs(y = 'Rain per event (mm)')

#response summary
rain_summary <- mrms_mi60 %>%
  ungroup() %>% 
  dplyr::select(c(site, P, MI60,
                  event)) %>%
  group_by(site) %>%
  summarize(Total_events = n(),
            Total_p_mm = sum(P),
            Max_P_mm = max(P),
            Mean_P_mm = mean(P),
            Max_MI60 = max(MI60),
            Mean_MI60 = mean(MI60)) %>%
  mutate_if(is.numeric,
            round,
            digits = 2)

kable(rain_summary) %>%
  row_spec(0,bold=TRUE) %>% 
    kable_styling()

```

# Stream response using MRMS

```{r stream response}

#get all stage data
load('stage.Rdata')

#tweak stage response function
get_radar_stage_resp <- function(events_df, stage_df) {
  
  response <- data.frame()
  
  for(i in 1:nrow(events_df)){
    temp <- stage_df %>%
      filter(datetime >=(events_df$starttime[i]) 
             & datetime <=(events_df$endtime[i] + 12*60*60))
    
    temp$event = ifelse(temp$datetime >=(events_df$starttime[i]) 
                        & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                        events_df$event[i])
    
    temp$duration_hr = ifelse(temp$datetime >=(events_df$starttime[i]) 
                              & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                              events_df$duration_hr[i])
    
    temp$P = ifelse(temp$datetime >=(events_df$starttime[i]) 
                    & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                    events_df$P[i])
    
    temp$MI60 = ifelse(temp$datetime >=(events_df$starttime[i]) 
                       & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                       events_df$MI60[i])
    
    temp$starttime = ifelse(temp$datetime >=(events_df$starttime[i]) 
                            & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                            events_df$starttime[i])
    
    temp$endtime = ifelse(temp$datetime >=(events_df$starttime[i]) 
                          & temp$datetime <=(events_df$endtime[i] + 12*60*60),
                          events_df$endtime[i])
    
    response <- rbind(response, temp)
  }
  
  out <- response %>%
    group_by(event) %>%
    mutate(Stage_range = max(Stage_mm) - min(Stage_mm)) %>%
    group_by(event) %>%
    filter(Stage_mm == max(Stage_mm)) %>%
    rename(Stage_mm_max = Stage_mm) %>% 
    group_by(event) %>%
    filter(datetime == max(datetime)) %>%
    rename(datetime_maxstage = datetime) %>%
    mutate(starttime_rain = as_datetime(starttime),
           endtime_rain = as_datetime(endtime),
           lag2p_hr = as.numeric(datetime_maxstage - starttime_rain)) %>%
    filter(!duration_hr == 0) %>%
    mutate(response = ifelse(Stage_range > 2, 1, 0)) %>% 
    dplyr::select(-c(starttime, endtime))
  
}


#break out mimrs_mi60 by site
me_events <- mrms_mi60 %>%
  filter(site == 'me')

mm_events <- mrms_mi60 %>%
  filter(site == 'mm')

mw_events <- mrms_mi60 %>%
  filter(site == 'mw')

ue_events <- mrms_mi60 %>%
  filter(site == 'ue')

um_events <- mrms_mi60 %>%
  filter(site == 'um')

uw_events <- mrms_mi60 %>%
  filter(site == 'uw')

#now get stage response for each site
me_response <- get_radar_stage_resp(me_events, me_stage)
mm_response <- get_radar_stage_resp(mm_events, mm_stage)
mw_response <- get_radar_stage_resp(mw_events, mw_stage)
ue_response <- get_radar_stage_resp(ue_events, ue_stage)
um_response <- get_radar_stage_resp(um_events, um_stage) 
uw_response <- get_radar_stage_resp(uw_events, uw_stage)

mrms_response <- bind_rows(me_response, mm_response,
                           mw_response, ue_response,
                           um_response, uw_response) %>%
  mutate(year = year(endtime_rain))

mrms_response %>%
ggplot(aes(site, Stage_range/10)) +
  geom_boxplot() +
  facet_wrap(~year) +
  ggtitle('Stage response amount') +
  labs(y = 'Stage response (cm)')

#response summary

response_summary <- mrms_response %>%
  ungroup() %>% 
  dplyr::select(c(site, P, MI60,
                  Stage_mm_max, Stage_range, lag2p_hr, response)) %>%
  group_by(site) %>%
  summarize(Total_events = n(),
            Total_pos_response = sum(response),
            Percent_response = (sum(response)/Total_events)*100,
            Max_Stage_Response_cm = max(Stage_range)/10,
            Mean_Stage_response_cm = mean(Stage_range)/10) %>%
  mutate_if(is.numeric,
            round,
            digits = 2)

kable(response_summary) %>%
  row_spec(0,bold=TRUE) %>% 
    kable_styling()


```

# Logistic regression 

Larimer County Bennett Creek streamflow gage, day of year, and rainfall metrics were used as predictors for response. 

```{r regress}

ben_q <- read_csv('/Users/megansears/Documents/Repos/CPF/bennett/larimerco_bennett_Q.csv') %>%
  rename(datetime = DateTime) %>%
  mutate(datetime = ymd_hm(datetime)) %>% 
  mutate(q_cms = q_cfs*0.028)

ben_q <- ben_q %>%
  rename(floor_dt = datetime)

#me 
me_response2 <- me_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
me_response2 <- left_join(me_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

me_glm <- glm(response ~ q_cms + doy + duration_hr + MI60 + P, data = me_response2, family = binomial)

me_response2$pred <- predict(me_glm, me_response2, type = 'response')

me_response3 <- me_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

me_kappa <- kappa2(me_response3)
#me_kappa$value

#mm
mm_response2 <- mm_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
mm_response2 <- left_join(mm_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

mm_glm <- glm(response ~ q_cms + doy + duration_hr + MI60 + P, data = mm_response2, family = binomial)

mm_response2$pred <- predict(mm_glm, mm_response2, type = 'response')

mm_response3 <- mm_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

mm_kappa <- kappa2(mm_response3)
#mm_kappa$value

#mw
mw_response2 <- mw_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
mw_response2 <- left_join(mw_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

mw_glm <- glm(response ~ q_cms + doy + duration_hr + MI60 + P, data = mw_response2, family = binomial)

mw_response2$pred <- predict(mw_glm, mw_response2, type = 'response')

mw_response3 <- mw_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

mw_kappa <- kappa2(mw_response3)
#mw_kappa$value

#ue
ue_response2 <- ue_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
ue_response2 <- left_join(ue_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

ue_glm <- glm(response ~ q_cms + doy + duration_hr + MI60 + P, data = ue_response2, family = binomial)

ue_response2$pred <- predict(ue_glm, ue_response2, type = 'response')

ue_response3 <- ue_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

ue_kappa <- kappa2(ue_response3)
#ue_kappa$value

#um
um_response2 <- um_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
um_response2 <- left_join(um_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

um_glm <- glm(response ~ q_cms + doy + duration_hr + MI60 + P, data = um_response2, family = binomial)

um_response2$pred <- predict(um_glm, um_response2, type = 'response')

um_response3 <- um_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

um_kappa <- kappa2(um_response3)
#um_kappa$value

#uw
uw_response2 <- uw_response %>%
  mutate(floor_dt = floor_date(datetime_maxstage, 'hour'))
    
uw_response2 <- left_join(uw_response2, ben_q, by = 'floor_dt') %>%
  mutate(doy = yday(floor_dt))

uw_glm <- glm(response ~ q_cms + doy + P + MI60 + duration_hr, data = uw_response2, family = binomial)
summary(uw_glm)

uw_response2$pred <- predict(uw_glm, uw_response2, type = 'response')

uw_response3 <- uw_response2 %>%
  mutate(prediction = if_else(pred > 0.5, 1, 0)) %>%
  ungroup() %>% 
  dplyr::select(response, prediction)

uw_kappa <- kappa2(uw_response3)
#uw_kappa$value

## make a df for this
first <- c('me', 'mm', 'mw',
           'ue', 'um', 'uw')
second <- c(me_kappa$value, mm_kappa$value, mw_kappa$value,
            ue_kappa$value, um_kappa$value, uw_kappa$value)

log_kappas <- data.frame(first,second) %>%
  rename(site = first,
         kappa = second)

kable(log_kappas) %>%
  row_spec(0,bold=TRUE) %>% 
  kable_styling()


```

# Next steps

* Continue developing logistic regression model (e.g., different models scenarios for TB and MRMS)
* Determine rain intensity thresholds
* Run comparison of stage change between sites (stage vs. intensity)
* Attributes of watersheds (slope, area, bedrock geology)
* Apply this workflow to East Troublesome


